%% Deadline for Full Papers: April 3, 2026, 11:59 pm Anywhere on Earth
%% The body of the paper should be a minimum of 5 pages and a maximum of 10 pages, with an unlimited number of pages allowed for references.
%% Reviewing will be non-blind; authors should include their names and contact information and reviews will include reviewer names.


%% Related work AI als nix besonders und Keine subsections sondern fett.

%% TODO: Make network one metric as the GMT can not distinquish bvetween bytes in and out

\documentclass[sigconf]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

\setcopyright{none}
\settopmatter{printacmref=false}
\acmDOI{}
\acmISBN{}
\acmConference[LIMITS '26]{12th Workshop on Computing within Limits}{June 23--25, 2026}{Online}

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{ifthen}
\usepackage{graphicx}
\usepackage{makecell}


% --- Listings setup (avoid minted for portability) ---
\lstdefinelanguage{json}{
  basicstyle=\ttfamily\footnotesize,
  showstringspaces=false,
  breaklines=true,
  morestring=[b]",
  literate=
   *{0}{{0}}{1}
    {1}{{1}}{1}
    {2}{{2}}{1}
    {3}{{3}}{1}
    {4}{{4}}{1}
    {5}{{5}}{1}
    {6}{{6}}{1}
    {7}{{7}}{1}
    {8}{{8}}{1}
    {9}{{9}}{1}
    {:}{{:}}{1}
    {,}{{,}}{1}
    {\{}{{\{}}{1}
    {\}}{{\}}}{1}
    {[}{{[}}{1}
    {]}{{]}}{1},
  morekeywords={true,false,null}
}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  columns=fullflexible,
  frame=single,
  rulecolor=\color{black!20},
  xleftmargin=0.0em,
  xrightmargin=0.0em,
  aboveskip=0.6em,
  belowskip=0.6em,
  escapeinside={(*@}{@*)}
}

\title{Measure Once, Model Everywhere:\\ Model-Based Per-Request Resource Consumption for HTTP}

\author{Geerd-Dietger Hoffmann}
\affiliation{
  \institution{Green Coding Solutions,\\ {\small University of Potsdam, HTW Berlin}}
  \city{Potsdam}
  \country{Germany}
}
\email{didi@green-coding.io}

\author{Verena Majuntke}
\affiliation{
  \institution{HTW Berlin}
  \city{Berlin}
  \country{Germany}}
\email{verena.majuntke@htw-berlin.de}

\begin{document}


\begin{abstract}
Recent proposals for HTTP-based sustainability disclosure focus on \textbf{what} environmental information should be transmitted at the protocol boundary, for example through response headers, but leave open the practical question of \textbf{how} such per-request values can be generated in realistic deployments. This paper addresses that implementation gap. We present a model-based approach for estimating resource consumption and \(\mathrm{CO_2e}\) per HTTP request without requiring fine-grained production power telemetry. The approach benchmarks endpoints offline under controlled conditions, derives compact endpoint-specific energy models from observable request features, and evaluates these models online at the HTTP server boundary. We implement this mechanism as an nginx extension that loads a JSON model registry and emits per-request metadata for energy, grid intensity, embodied emissions, and total request-level impact. Using a benchmark application with conventional CRUD (Create, Read, Update, Delete) endpoints and an LLM inference endpoint, we show that heterogeneous request classes can be represented with constant, linear, and piecewise models, including token-based modelling for inference workloads. Our evaluation indicates that the approach is operationally feasible and introduces only low runtime overhead.
\end{abstract}
\maketitle

%-----------------------------------------------------------------------------------------
\section{Introduction}
%-----------------------------------------------------------------------------------------

Software is increasingly delivered through HTTP interfaces built on opaque
stacks of virtualised infrastructure, managed platforms, and third-party
services. While this architecture improves scalability and development
velocity, it obscures the environmental impact of individual actions
\cite{parssinen2018onlineadvertising,han2025fairco2}. Users can observe latency
and price, but not the energy use or emissions of a single request. Providers
face a similar challenge: even when disclosure is desired, request execution
spans multiple systems, and fine-grained power telemetry is typically
unavailable.

A recurring principle in sustainable computing is to expose information at the
interface where decisions are made. The Software Carbon Intensity (SCI)
specification expresses emissions relative to a functional unit \(R\)
\cite{gsf-sci}. For web systems, the natural functional unit is the HTTP
request. If request-level impact were available at the protocol boundary,
clients could incorporate it into decisions such as retry behaviour, batching,
or prompt design.

HTTP is a natural disclosure surface, and prior work has proposed response headers for communicating carbon information~\cite{martin-http-carbon-emissions-scope-2-00,httpwg-admin-issue-52-carbon-emissions,ietf-http-wg-mailinglist-carbon-header-thread,besleaga-green-sustainability-header-00}. However, these efforts focus on \emph{what} to disclose, not on how to derive per-request values in practice.

From a resource-reduction perspective, however, disclosure is valuable
only insofar as it supports transitions toward operating within
ecological constraints rather than optimising the status quo more
precisely.  Nardi et al.\ argue that computing research should make
growth-oriented assumptions visible and contestable~\cite{nardi2018computingwithinlimits}.
Per-request sustainability metadata serves this function in a specific
way: it makes the marginal environmental cost of software demand
legible at the exact interface where that demand is created.  Without
such legibility, neither providers nor consumers can distinguish
high-cost from low-cost interactions, and sustainability
considerations remain structurally excluded from the design decisions
that shape aggregate resource consumption.  The mechanism is therefore
not an end in itself, but a precondition for constraint-aware software
practice.

In production systems, per-request energy and emissions are not directly observable. Attribution is complicated by multi-tenancy, virtualization, and
request fan-out, while hardware telemetry is coarse or unavailable. The core challenge is therefore not the disclosure interface, but the construction of a
method that produces meaningful per-request estimates under realistic constraints.

This paper addresses this gap by introducing a practical, model-based method for deriving per-request energy and emissions. Instead of relying on production telemetry, endpoints are benchmarked offline under controlled conditions to derive compact energy models. These models are evaluated at request time using variables already observable at the HTTP boundary, such as request duration, transferred bytes, or application-level metadata. HTTP-based disclosure is used as one instantiation, but is not required by the method.

This paper makes four contributions:
(1) It separates the protocol-level question of \emph{what to disclose} from the systems question of \emph{how to generate request-level values}, and
identifies the latter as a key gap in current HTTP sustainability discussions.
(2) It proposes a measurement-to-model pipeline that derives compact, interpretable, and executable energy models for HTTP endpoints from controlled
offline benchmarks.
(3) It demonstrates how these models can be evaluated at the HTTP server boundary and exposed as request-level sustainability metadata, while keeping
underlying assumptions explicit.
(4) It evaluates the feasibility, expressiveness, and runtime overhead of the approach across heterogeneous endpoints, including conventional web services
and AI inference workloads.


%%%%%%%%%%%%
\section{Related Work}
\label{sec:rw}
%%%%%%%%%%%%%

Assessing the environmental impact of software requires more than raw
instrumentation. Guldner et al.\ propose the Green Software Measurement Model
(GSMM), which emphasises that sustainability assessment depends on explicit
modelling choices, system boundaries, and the interpretation of measurements
\cite{guldner2024gsmm}. The Software Carbon Intensity (SCI) specification
similarly frames impact relative to a functional unit \(R\) \cite{gsf-sci}.
Together, these approaches highlight that measurement alone is insufficient and
must be complemented by modelling.

Prior work also shows that energy use varies significantly with workload
characteristics. Pärssinen et al.\ quantify the environmental impact of web
interactions \cite{parssinen2018onlineadvertising}, and Sala et al.\ demonstrate
that sustainability metrics can be integrated into operational web monitoring
\cite{sala2024greenwebmeter}. In the context of AI, Desislavov et al.\ and
Poddar et al.\ show that inference energy depends on runtime properties such as
execution time and output length \cite{desislavov2023trends,poddar2025towards}.
These works establish that fine-grained variation exists, but primarily analyse
it offline or at aggregate levels.

Attributing energy and emissions in cloud environments is inherently difficult.
Fair-CO2 highlights the challenge of allocating both operational and embodied
emissions across shared infrastructure and multi-tenant systems
\cite{han2025fairco2}. This limits the availability of directly observable
per-request values in production environments.

Recent work has explored HTTP as a surface for communicating sustainability
information. Martin proposes a "Carbon-Emissions-Scope-2" response header
\cite{martin-http-carbon-emissions-scope-2-00}, and more recent drafts define
structured sustainability headers \cite{besleaga-green-sustainability-header-00}.

Existing work therefore establishes the need for modelling, demonstrates
workload-level variation, and defines potential disclosure interfaces.
However, it does not provide a practical method for generating per-request
energy and emissions estimates under realistic deployment constraints. In
particular, the step from controlled measurements to request-level values
available at the interface remains insufficiently specified.


%-----------------------------------------------------------------------------------------
\section{The Measurement Gap}
\label{sec:gap}
%-----------------------------------------------------------------------------------------
Software decisions in HTTP-based systems are made at the level of individual
requests. Requests are the unit at which services are invoked, billed,
rate-limited, and experienced, and the unit at which clients can adapt
behaviour, for example by retrying later, batching work, reducing payload size,
or choosing a different endpoint or provider. In the SCI sense, the HTTP
request is therefore a natural functional unit for many web-facing systems
\cite{gsf-sci}.

At the same time, the environmental impact of a single request is not directly
observable in production environments. Power telemetry, where it exists, is
typically available only at machine, rack, or datacentre level rather than at
transaction level. Individual requests execute concurrently, share compute,
memory, storage, and network resources, and often trigger additional work in
caches, databases, queues, or external services. In cloud and platform
environments, virtualization and multi-tenancy further complicate attribution.
As a result, servers can observe request-level metadata such as duration,
transferred bytes, or status codes, but not the physical energy consumption of
individual requests.

This creates a structural mismatch at the interface where decisions are made.
Clients can already observe latency, quotas, and price, but they typically
cannot observe the environmental cost of a request. This limits their ability
to incorporate sustainability into concrete software decisions such as caching,
batching, retry strategies, response-size control, or prompt reduction.
Providers face the complementary problem: even when they want to analyse or
disclose impact, they lack a practical way to derive request-level values from
the infrastructure they operate. In other words, the disclosure surface exists
at the HTTP boundary, but the quantity to disclose is not readily available
there.

Given these constraints, per-request energy cannot be obtained through direct
measurement in typical production environments. Fine-grained power telemetry
is rarely available, and even in controlled settings, machine-level
measurements require careful baseline handling, repetition, and explicit
allocation choices to produce stable results. Measurement alone therefore does
not yield request-level values. A practical solution requires a method that transforms controlled measurements
into estimates that can be computed at request time from variables already
observable at the interface.

Such a method must satisfy a set of design requirements derived based on the system context. First, it must operate
under limited observability, including deployments without fine-grained power
sensors. Second, it must be efficient enough to evaluate on every request at
the server boundary. Third, it must remain interpretable, so that providers can
understand and communicate the assumptions and system boundaries underlying the
reported values. Fourth, it must capture heterogeneous endpoint behaviour,
ranging from near-constant-cost CRUD requests to highly variable AI inference
workloads whose cost depends on request-specific inputs.

We address these requirements with a model-based approach. Endpoints are
benchmarked offline under controlled conditions to derive compact,
endpoint-specific energy models. At request time, these models are evaluated
using variables that are already observable at the HTTP boundary, such as
request duration, transferred bytes, method, status, and selected request
parameters. Where generic HTTP features are insufficient, applications can
expose additional request-scoped variables, such as token counts for inference
workloads. The resulting per-request energy estimates are then combined with
grid-intensity and embodied-allocation parameters to derive request-level
\(\mathrm{CO_2e}\) values.


%-----------------------------------------------------------------------------------------
\section{Per-Request Energy Estimation}
\label{sec:methods}
%-----------------------------------------------------------------------------------------

This section presents our approach for deriving per-request energy and
emissions estimates from controlled measurements under realistic observability
constraints. The approach separates offline measurement from online estimation.
Endpoints are first characterised under controlled conditions, and the
resulting models are evaluated at request time using variables observable at
the HTTP boundary.

We first define endpoint-level energy models based on runtime-observable
features such as request duration, transferred bytes, and request-specific
parameters in Section~\ref{subsec:model}. We then describe how predicted
energy values are translated into operational and embodied emissions using
grid-intensity and allocation parameters in
Section~\ref{subsec:emissions}. Finally, we discuss the limitations of this
approach, including its dependence on modelling assumptions and
deployment-specific calibration in Section~\ref{subsec:limitations}.


\subsection{Endpoint energy models}
\label{subsec:model}

We model the energy consumption of HTTP endpoints as a function of
runtime-observable request features. These include, for example, request
duration, transferred bytes, HTTP method, response status, and
request-specific parameters. The set of features is not fixed. Rather, it
captures quantities that are either already available at the HTTP boundary or
can be exposed with minimal application changes. In practice, such features
reflect the dominant sources of variation in per-request energy use, including
compute, data transfer, and input-dependent workload characteristics.

Formally, let \(e\) denote an endpoint and \(x\) a vector of selected features.
We fit a model \(\widehat{E}_e(x)\) that predicts the energy per request (in
\(\mathrm{mWh}\)). To obtain models that are both practical and robust, we
restrict the model space to a small set of families that balance
expressiveness, interpretability, and efficient runtime evaluation. The goal
is not to capture arbitrary complexity, but to cover common patterns of
endpoint behaviour with models that remain simple enough to evaluate at the
HTTP boundary.

In practice, we observe three recurring patterns. Some endpoints exhibit
negligible variation across requests, others scale approximately with
observable request features such as processing time or payload size, and some
show non-linear or threshold behaviour. We therefore use three model families
that reflect these cases.

\textbf{Constant model:}
For endpoints whose energy is stable across input variation, we use a constant
model:
\[
\widehat{E}_e = c_e.
\]

\textbf{Linear model:}
For endpoints whose energy scales with observable request features, we use a
linear model. Let \(x = (x_1,\dots,x_p)\) denote the selected feature vector.
We then predict per-request energy as
\[
\widehat{E}_e(x) = \beta_0 + \sum_{j=1}^{p} \beta_j x_j.
\]
That is, the predicted energy is modelled as a constant overhead term
\(\beta_0\) plus the weighted sum of the selected request features, where each
coefficient \(\beta_j\) captures the contribution of feature \(x_j\) to the
endpoint's energy consumption.

\textbf{Curve model:}
For endpoints whose energy does not scale well under a single global linear
relationship, we use a piecewise linear model. As with the linear model, the
prediction is based on selected request features. In this case, we choose a
single scalar feature \(z\), for example request size, response size, or a
numeric application-level parameter, and define \(n\) measured support points
\[
\{(z_i, E_i)\}_{i=1}^{n},
\]
where \(z_i\) is the value of the selected feature and \(E_i\) the
corresponding per-request energy. For an observed value \(z\), the predicted
energy is obtained by linear interpolation between adjacent support points:
\[
\widehat{E}_e(z)
=
E_i + \frac{z-z_i}{z_{i+1}-z_i}(E_{i+1}-E_i),
\qquad z_i \leq z \leq z_{i+1}.
\]
Thus, the endpoint energy is represented as a sequence of linear segments over
the measured range of the selected feature, allowing the model to capture
changes in scaling behaviour while remaining simple and efficient to evaluate
at request time.

\textbf{Application-supplied model inputs:}
Some relevant predictors are not observable at the HTTP server layer, but are
known to the application or to a downstream runtime. In such cases, the
application can expose additional request-scoped variables that can be used as
inputs to the same modelling approach described above. These variables may
include, for example, token counts for \texttt{LLM}-style inference, result-set
cardinality, number of database records touched, cache-hit indicators from the
application layer, prompt length, or other domain-specific workload
descriptors.

Formally, such variables can be treated as an extension of the
runtime-observable feature vector \(x\), including both server-observable
features and application-supplied inputs. The general modelling approach
remains the same: per-request energy is predicted from request-scoped
variables available at evaluation time.

As a concrete example, for an inference endpoint, token-based variables can be
included as part of the feature vector, yielding a specialised model such as
\[
\widehat{E}_{ai}(\tau^{prompt}, \tau^{gen})
=
\alpha_0 + \alpha_p \tau^{prompt} + \alpha_g \tau^{gen}.
\]
where \(\tau^{prompt}\) denotes the number of prompt tokens and
\(\tau^{gen}\) the number of generated tokens.
This reflects the empirical observation that, under fixed configuration (e.g.,
temperature, context length, batch size), the energy consumption of many
decoder-only models can be approximated from token-related workload variables.

More generally, application-level features can be incorporated whenever they
improve predictive accuracy and can be exposed at request time. Token counts
are one instance of this pattern; other domain-specific variables may be used
in the same way, provided they can be made available to the reverse proxy as
request-scoped metadata (e.g., via internal headers or server variables).

\textbf{Registry generation procedure.}
In the current prototype, model-family assignment is based on
endpoint-specific workload-scaling benchmarks rather than on an automatic
search over candidate families. For each endpoint, we constructed a benchmark
that progressively increased the relevant workload and measured how the energy
profile changed. For example, for \texttt{/login} we issued between 1 and
1000 calls in order to determine whether per-request energy remained
stable or varied with increasing load\footnote{\url{https://metrics.green-coding.io/stats.html?id=6c40a5b8-79b7-4372-a1c4-3ae7185451d2}}.
We applied the same procedure to all endpoints and then selected the model family that best matched the observed
behavior. Concretely, \texttt{/login}, \texttt{/logout}, \texttt{/done}, and
\texttt{/deleteAllToDos} are represented as constant models whose value is the
mean measured per-request energy across their benchmark conditions.
\texttt{/createToDo} and \texttt{/getToDos} are represented as curve models:
the measured support points are retained, duplicate \(x\)-values are averaged,
and runtime estimates are obtained by piecewise linear interpolation between
the resulting points. For \texttt{/ai}, we fit a least-squares linear model
over an intercept plus prompt-token and generated-token counts. This
benchmark-driven route-to-family mapping keeps the generated registry simple
and inspectable, but it also means that the current prototype does not yet
perform automatic family selection or uncertainty-aware fitting. Future work
will include a script that automatically generates a model registry from GMT benchmark data.


\subsection{Operational and embodied emissions}
\label{subsec:emissions}

The models above provide per-request energy estimates. To obtain a more interpretable sustainability metric, we convert predicted energy into carbon emissions by accounting for both operational energy use and embodied hardware impact.

In the prototype implementation, per-request carbon outputs are emitted in \(\mathrm{mgCO_2e}\). We therefore separate operational and embodied components in that unit. This distinction captures both the energy consumed during request execution and the amortised impact of the underlying hardware.

\textbf{Operational emissions.}
Operational emissions are derived directly from the predicted energy consumption. Let \(\widehat{E}_e\) denote the predicted energy per request (in \(\mathrm{mWh}\)) and \(I\) the grid carbon intensity (in \(\mathrm{gCO_2e/kWh}\)). Since \(\SI{1}{kWh} = 10^6\,\si{mWh}\) and \(\SI{1}{g} = 1000\,\si{mg}\), the operational emissions emitted by the prototype are:
\[
\widehat{C}^{op}_e = \widehat{E}_e \cdot \frac{I}{1000},
\]
with \(\widehat{C}^{op}_e\) expressed in \(\mathrm{mgCO_2e}\).

\textbf{Embodied emissions as a time-based allocation.}
Embodied emissions capture the carbon cost of manufacturing and provisioning the hardware. Let \(M\) denote the total embodied emissions of the system (in \(\mathrm{gCO_2e}\)), and \(T\) an assumed operational lifetime (in seconds). We model embodied emissions as a constant emission rate stored in the registry as:
\[
r^{emb} = \frac{M}{T}\ \mathrm{\Bigl(\frac{gCO_2e}{s}\Bigr)}.
\]
For a request with processing time \(t\), the prototype converts this rate to \(\mathrm{mgCO_2e}\) and emits:
\[
\widehat{C}^{emb}_e = 1000 \cdot r^{emb} \cdot t,
\]
with \(\widehat{C}^{emb}_e\) expressed in \(\mathrm{mgCO_2e}\).

Finally, total per-request emissions are defined as the sum of operational and embodied components:
\[
\widehat{C}_e = \widehat{C}^{op}_e + \widehat{C}^{emb}_e.
\]

The prototype outputs should be read under four explicit accounting
assumptions. First, benchmark labels are average per-request values obtained by
dividing measured machine energy by the number of homogeneous requests in a
benchmark block. Second, the registry used in this paper is derived from total
machine energy rather than idle-subtracted energy. Third, HTTPS/TLS overhead is
excluded from the current calibration. Fourth, embodied carbon is allocated as
a constant time-based rate. The reported numbers are therefore
deployment-specific estimates under a stated accounting model rather than
protocol-independent constants.

This formulation provides a simple and transparent allocation scheme. Alternative approaches, such as utilisation-weighted or capacity-based allocation, can be substituted depending on the deployment context \cite{iso14040,iso14044,ghgproduct,gsf-sci}.

\subsection{Limitations}
\label{subsec:limitations}
However, the approach is subject to limitations. Per-request energy and
emission estimates depend on deployment-specific factors such as hardware
generation, power-management settings, co-located workloads, virtualization
layers, and regional infrastructure. As a result, the same logical endpoint
may exhibit different physical footprints across environments. The estimates
produced by our model should therefore be understood as interface-level
abstractions derived from controlled measurements, explicit assumptions, and
runtime-observable features. They are not intended to provide exact physical
attribution for individual requests, but to expose a consistent and actionable
signal at the boundary where software decisions are made.

A further limitation follows from the dependence on the measurement setup. If
the deployment environment differs significantly from the calibration
environment, prediction error may increase and deployment-specific effects may
not be fully captured. In addition, the model reflects only those factors that
are observable or explicitly included in the feature set. Unobserved effects,
such as resource contention, background load, or hardware-specific
optimisations, may influence energy consumption but are not represented, which
can introduce systematic error.

The current prototype also selects model
families from endpoint-specific workload benchmarks rather than by automatic
model selection, and it does not yet emit uncertainty intervals alongside
point estimates. The approach therefore abstracts over dynamic system
behaviour, so variations in load, scheduling, or co-tenancy may lead to
request-level fluctuations that are not captured even when the feature inputs
remain unchanged.

Finally, as workloads, system configurations, or
infrastructure evolve over time, model parameters may become less accurate,
making periodic recalibration necessary. It is therefore necessary to automate this step so that the model registry can be updated on new software releases or hardware changes.

Despite these limitations, the approach provides a structured and reproducible abstraction with explicitly defined inputs, assumptions, and system boundaries. The resulting estimates should therefore be interpreted as context-dependent and primarily suited for relative comparison and decision support at the software interface.


%-----------------------------------------------------------------------------------------
\section{Measurement Setup and Data Collection}
%-----------------------------------------------------------------------------------------
\label{sec:measurement}
This section describes how we generate per-request energy labels \(E\) (in \(\mathrm{mWh}\)) from controlled experiments. These measurements form the basis for fitting the per-endpoint energy models described in Section~\ref{sec:methods}. For each request, we record energy at the machine boundary and associate it with request-level observations obtained during benchmarking, including request processing time, data transfer volume, and endpoint-specific workload parameters.

All measurements were executed using the Green Metrics Tool (GMT) \cite{greenmetricstool} in a controlled measurement setup. Experiments were conducted on a dedicated machine to ensure reproducibility and minimise interference from background workloads.

We used the GMT ``ML/AI Profiling'' profile on a server-class system with the following configuration:
\begin{itemize}
  \item CPU: Intel Core i5-9600K @ 3.70\,GHz (6 cores / 6 threads; Turbo enabled; DVFS enabled)
  \item Memory: 32\,GB
  \item GPU: NVIDIA GeForce GTX 1080 (8\,GB VRAM), CUDA 12.2, driver 535.274.02
  \item Power measurement: MCP39F511N \& IPMI (as provided by GMT)
  \item OS: Ubuntu 24.04 (NOP Linux) \cite{greenmetricstool}
\end{itemize}

Turbo boost and dynamic frequency scaling were left enabled to reflect typical deployment conditions.
\subsection{Measurement Practice}
To obtain stable energy measurements, we followed the Green Coding best practices for controlled benchmarking \cite{greencoding_bestpractices}. These practices stress that software-energy measurements are sensitive to noise and should therefore be performed under controlled conditions, with temperature effects, idle baselines, and run-to-run variance taken into account.

Concretely, we ran the service under test (SUT) on an isolated machine running NOP Linux, a minimal environment with non-essential processes disabled to reduce measurement noise. Before sampling, we warmed up the SUT to reach a steady state. We then recorded an idle baseline for a comparable interval. During the experiments, we varied one workload factor at a time, such as payload size or token count, and repeated each condition often enough to make the signal distinguishable from measurement noise. As recommended in the Green Coding documentation, the exact number of repetitions depended on the variance and expected signal strength of the endpoint under test.

The whole test suite was executed 10 times. All measurements are available under \url{https://tinyurl.com/fprne67y}.

\subsection{Benchmark Application Setup}
We implemented a minimal REST-style ToDo application in Python that serves HTTP
requests using SQLite as the backend database. A user can log in and log out,
create ToDo items with a title, text, and optional file attachment, retrieve
all ToDo items, and mark individual items as done. In addition, the application
provides an LLM endpoint for sentence completion. The application and benchmark
harness are publicly available \cite{easytodo}. While SQLite is a simple
dependency, the GMT harness supports measuring more complex multi-service
deployments. We use this application to cover a range of endpoint behaviours.
The endpoints are described in Table~\ref{tab:endpoints}.

\begin{table}
    \centering
    \begin{tabular}{p{1.7cm} p{6.3cm}}
        /login & Given a username and password returns an authenticated session cookie \\
        /logout & Invalidates the authenticated session \\
        /createToDo  & Takes a title, text and attachment and creates a new "todo" item in the database \\
        /done &  Given a todo id marks it as done in the DB \\
        /getToDos  & Returns all the todos of the user \\
        /deleteAllToDos & Deletes all the todos of the user \\
        /ai & Given a text sends it to an LLM for autocompletion and returns the ai response \\
    \end{tabular}
    \caption{REST endpoints used in the benchmark application}
    \label{tab:endpoints}
\end{table}

We deliberately excluded HTTPS from the measurements to avoid entangling endpoint energy with TLS handshake and encryption overhead. In deployments where HTTPS is mandatory (typical on the public internet), TLS overhead can be accounted for as (i) a separate endpoint-independent constant term, or (ii) a distinct model variant keyed by protocol configuration.


\subsection{Benchmarking}
As introduced in Section~\ref{sec:methods}, the prototype uses constant,
linear, and curve models. Each endpoint was benchmarked with its own workload
script designed to expose the endpoint's expected scaling factors. Stateful
endpoints requiring authentication were benchmarked with a pre-acquired session
cookie. The cookie was captured once and reused across runs to avoid repeatedly
measuring login overhead when benchmarking other endpoints.

We do not use one uniform repetition count for all endpoints. Instead, we choose the number of calls per benchmark condition according to the expected signal strength. For very small endpoints such as \texttt{/login} and \texttt{/logout}, single-request energy values are too small relative to machine-level measurement noise, so we execute 100 calls per condition and divide the measured machine energy by the number of calls. We use the same 100-call batch size for the CRUD-style web endpoints \texttt{/createToDo} and \texttt{/getToDos}. For \texttt{/done}, each benchmark condition first creates 100 ToDo items and then marks each of them once, yielding 100 \texttt{/done} requests for that condition. For \texttt{/ai}, we use 10 requests per prompt profile because each inference call is much longer and already produces a clearly measurable energy signal.

For stateful workload blocks we keep the batching strategy aligned with the application semantics. A \texttt{/createToDo} condition creates 100 items with either a fixed text length (100, 1000, 10000, or 100000 characters) or a fixed attachment size (1\,KB, 10\,KB, 100\,KB, 1\,MB, or 5\,MB). The subsequent \texttt{/getToDos} condition is then executed 100 times against that populated state. The \texttt{/done} condition marks the 100 created items exactly once each, which again yields 100 measured requests. After each block, \texttt{/deleteAllToDos} is called only for cleanup so that the next condition starts from a controlled and comparable state. In the AI case, the benchmark uses five prompt profiles and issues 10 requests per profile, because inference latency and energy are high enough that additional repetition is less critical for signal extraction than for the lightweight REST endpoints.

For the \texttt{/ai} endpoint we used an \texttt{ollama} container to serve local inference on the same machine. We used \texttt{gemma3:1b}, a lightweight text-only model from Google's Gemma 3 family, as the inference backend for our experiments. Its comparatively small parameter size makes it suitable for local serving scenarios while still supporting context windows of up to 32k tokens. To stabilise inference behaviour:
\begin{itemize}
  \item the model was downloaded before the benchmark window started,
  \item caches were prewarmed (warm-up requests executed before sampling),
  \item the inference configuration was held constant across runs (e.g., decoding settings), so that token count is an informative proxy for work.
\end{itemize}

We treat token counts as first-class measurement variables for \texttt{/ai}. In our prototype, the application layer observes both prompt-token and generated-token counts via the inference runtime and exposes them as response-scoped metadata. This enables fitting and evaluating token-linear models \(\widehat{E}_{ai}(\tau)\) without requiring invasive GPU telemetry in production.

Prior work suggests that generated token count is a practical proxy for LLM inference cost under fixed serving conditions. Output length has been shown to scale approximately linearly with inference latency, and prior energy studies also report strong correlations between energy use and output-token count \cite{yang2024queueing,poddar2025towards,caravaca2025promptstopower}. This makes it feasible to use generated token count as a factor in our \texttt{/ai} endpoint model.

%\subsection{Deriving per-request energy labels from GMT exports}
\subsection{Per-Request Energy Labels}

We derive per-request energy labels \(E\) (in \(\mathrm{mWh}\)) from machine-level energy measurements collected during benchmarking\footnote{\url{https://tinyurl.com/mryh2wm8}}. For each condition, we compute:
\[
E = \frac{E_{\text{machine}}}{n},
\]
where \(E_{\text{machine}}\) is the total energy measured during the benchmark interval and \(n\) is the number of requests executed in that interval.

This yields an average per-request estimate that approximates the energy attributable to the endpoint under the given workload condition. The estimate assumes that requests within each condition are homogeneous and that background effects remain stable during the measurement interval.

We verify that energy varies consistently with the controlled workload parameter under input sweeps, indicating that the derived labels capture the intended endpoint behaviour.

\subsection{Runtime-Observable Features}
For each measured request, we associate the energy label \(E\) with features that are available (or can be made available at low cost) at the HTTP server boundary:
\begin{itemize}
  \item \textbf{Processing time \(t\):} request handling time as observed by the reverse proxy (e.g., nginx \texttt{\$request\_time}-style timing variables).
  \item \textbf{Request and Response bytes \(b^{req}\), \(b^{res}\):} the size of incoming and outgoing data, capturing transfer-related effects.
  \item \textbf{HTTP method \(m\):} the request method (e.g., \texttt{GET}, \texttt{POST}, \texttt{PUT}, \texttt{DELETE}), which helps distinguish different execution paths and backend behaviours.
  \item \textbf{Response status \(r\):} captures outcomes such as success, redirection, rejection, or failure, which may correlate with different execution costs.
  \item \textbf{URL-parameter-derived features \(q\):} selected information derived from the request URI, such as the presence of specific query parameters, total query-string length, parameter count, or whitelisted numeric values (e.g., pagination limit). We use derived features rather than raw query strings in order to avoid unnecessary sparsity and high-cardinality model inputs.
  \item \textbf{Token count \(\tau\):} for AI inference, prompt and generated token counts exposed by the application.
\end{itemize}

The resulting dataset is a collection of tuples \((e, x, E)\) per endpoint \(e\), where \(x\) is the vector of relevant features. We then fit the constant, linear, and piecewise curve models described in Section~\ref{sec:methods}, and export the fitted parameters as will be described in the next section.


\section{HTTP Server Implementation}
\label{sec:implementation}
This section describes how the proposed model-based approach is realised at the HTTP server boundary. We implement per-request energy estimation and disclosure as an extension to the web server, enabling models derived from offline benchmarking to be evaluated at request time. The implementation focuses on integrating model evaluation into the request-processing path with minimal overhead, and on exposing the resulting energy and emissions values through standard HTTP mechanisms.

\subsection{Design and Runtime Behaviour}
The implementation realises per-request energy estimation at the HTTP server layer using an nginx extension. At startup, the module loads a JSON registry that maps URI paths (or location blocks) to endpoint-specific energy models and associated parameters. These models are derived offline and encoded in a compact form that can be evaluated efficiently during request processing. At runtime, the module operates as part of the response handling pipeline. For each request, it first resolves the matching path and then, where configured, applies additional filter criteria such as the HTTP method to select the appropriate model variant. It then extracts the required input features from variables already available in nginx, including request processing time, request and response sizes, and, where applicable, additional request-scoped metadata such as token counts provided by the application.

To minimise overhead, the registry is normalised once at worker initialisation. Route definitions are canonicalised, model coefficients are preloaded, and curve support points are parsed and sorted. Model evaluation is implemented as a small deterministic computation. Constant models return a fixed value, linear models apply preconfigured coefficients to the observed features, and piecewise models perform interpolation over precomputed support points. The resulting energy estimate is then converted into operational and embodied emissions using the configured grid intensity and embodied allocation parameters. The computed values are exposed at the HTTP boundary through response headers and can optionally be written to structured logs for downstream analysis. The implementation relies on the \texttt{ngx\_http\_js\_module} using the \texttt{njs} runtime, which allows extending nginx without requiring kernel modifications or external telemetry interfaces.

\subsection{Model Representation}
The JSON registry defines how endpoint-specific energy models and their parameters are represented for runtime evaluation. At the top level, the registry is a map from endpoint identifiers (e.g., URI paths or location blocks) to per-endpoint configuration objects. Each such object contains an \texttt{energy\_model} plus the shared parameters needed later for emission calculations, namely grid intensity and the embodied-allocation rate. The \texttt{energy\_model} itself is a tagged structure whose fields depend on the chosen model family: constant models store a single value, linear models store an intercept and coefficients, and curve models store support points together with interpolation rules. The registry format also allows multiple model variants for the same URL path, distinguished by an optional \texttt{filter} object. This is useful when, for example, \texttt{GET} and \texttt{POST} requests share the same endpoint path but exhibit different behaviour and should therefore be matched to different energy models. Listing~\ref{lst:json-registry} does not show the full generated registry, but a compact excerpt with representative instances of the model families used in the prototype.


\begin{lstlisting}[float,language=json,caption={Excerpt from the JSON model registry with representative endpoint entries},label={lst:json-registry}]
{
  "/ai": {
    "filter": { "method": "POST" },
    "energy_model": {
      "kind": "linear",
      "intercept_mWh": 7.062,
      "prompt_token_coeff_mWh_per_token": 0.017,
      "generated_token_coeff_mWh_per_token": 0.483
    },
    "embodied_rate_gCO2e_per_s": 0.001,
    "grid_intensity_gCO2e_per_kWh": 121.111
  },
  "/done": {
    "energy_model": {
      "kind": "constant",
      "value_mWh": 0.085
    },
    "embodied_rate_gCO2e_per_s": 0.001,
    "grid_intensity_gCO2e_per_kWh": 121.111
  },
  "/getToDos": {
    "energy_model": {
      "kind": "curve",
      "input": "data_size",
      "points": [[49421.59, 0.072], [231310.93, 0.112], [2051378.03, 0.125], [20298823.13, 0.553]],
      "interpolate": "linear",
      "extrapolate": "linear_tail"
    },
    "embodied_rate_gCO2e_per_s": 0.001,
    "grid_intensity_gCO2e_per_kWh": 121.111
  },
  ...
}
\end{lstlisting}

Energy values are expressed in \(\mathrm{mWh}\), which provides sufficient precision for request-level estimates while avoiding numerical underflow. Grid intensity is represented in \(\mathrm{gCO_2e/kWh}\), and embodied emissions are encoded as a time-based rate in \(\mathrm{gCO_2e/s}\), allowing per-request embodied impact to be derived from request duration. Separating energy models from their execution environment allows the same model structure to be reused across deployments while adapting parameter values to local conditions. In particular, grid intensity is treated as a configurable input rather than a fixed constant. Depending on the deployment, it may be defined statically, updated periodically, or obtained from an internal service. This representation makes the assumptions underlying the reported values explicit.


\subsection{Disclosure Interface}
\label{subsec:interface}
The computed energy and emissions values are exposed at the HTTP boundary using response headers. This interface makes per-request sustainability information available to clients at the point where software decisions are made, without requiring changes to the application protocol or additional service endpoints. Rather than encoding all information in a single aggregated value, we expose multiple fields that separate the underlying components of the calculation. This includes request-level energy, grid intensity, and the resulting operational and embodied emissions. Keeping these quantities separate makes the assumptions of the model explicit and allows clients and operators to interpret, compare, and recombine the values as needed. The following response headers are used in the prototype implementation:
\begin{itemize}
  \item \texttt{Request-Energy}: Energy attributable to handling the request, in \(\mathrm{mWh}\).
  \item \texttt{Grid-Intensity}: Grid intensity used for operational conversion, in \(\mathrm{gCO_2e/kWh}\).
  \item \texttt{Request-Embodied-CO2e}: Embodied emissions allocated to the request, in \(\mathrm{mgCO_2e}\).
  \item \texttt{Request-SCI}: Total emissions for the request, in \(\mathrm{mgCO_2e}\).
\end{itemize}
These fields represent one possible design for exposing per-request sustainability information and are not intended as a proposal for standardisation. The separation of components is chosen to make modelling assumptions explicit during experimentation and evaluation, rather than to prescribe a fixed interface. Listing~\ref{lst:curl} shows a \texttt{curl}-based \texttt{POST} request to the \texttt{/login} endpoint together with the corresponding HTTP response produced by the prototype \texttt{nginx} module. This listing includes the four sustainability-related headers introduced above. We have removed data that is not relevant for this example like the date, keep-alive information, etc. This illustrates how per-request energy and carbon information is disclosed at the HTTP boundary in a concrete interaction.

\begin{lstlisting}[float,language=bash,caption={Example response headers emitted by the prototype nginx module},label={lst:curl}]
$ curl -sS -D - -o /dev/null \
  -X POST "http://localhost:8080/login" \
  -H "Content-Type: application/json" \
  -d "{\"username\":\"$USERNAME\",\"password\":\"$PASSWORD\"}"

HTTP/1.1 200 OK
Server: nginx/1.25.5
...
Set-Cookie: csrftoken=xxx; ...
Set-Cookie: sessionid=xxx; ..
(*@{\bfseries\ttfamily Request-Energy: 4.038}@*)
(*@{\bfseries\ttfamily Grid-Intensity: 121.111}@*)
(*@{\bfseries\ttfamily Request-Embodied-CO2e: 0.143}@*)
(*@{\bfseries\ttfamily Request-SCI: 0.633}@*)
\end{lstlisting}



\section{Evaluation}
\label{sec:evaluation}

We evaluate the proposed approach using the controlled benchmark setup described in Section~\ref{sec:measurement} and the HTTP-based implementation introduced in Section~\ref{sec:implementation}. The evaluation focuses on whether per-request energy models derived from machine-level measurements can be constructed, interpreted, and executed at the HTTP boundary under realistic conditions. In particular, we examine whether benchmark measurements can be converted into a complete model registry across heterogeneous endpoints, whether the fitted models preserve expected workload-dependent behaviour for both CRUD-style and AI-style requests, and whether the resulting disclosure mechanism remains lightweight enough for reverse-proxy deployment.

The evaluation is intentionally framed as a feasibility study rather than a
full predictive-accuracy benchmark. The benchmark suite was executed ten times
to stabilise the underlying measurements, after which we inspect the derived
model parameters, perform one trace-based replay validation against an
independent click-through workload, and compare paired runs with and without
header emission for overhead. The results therefore support claims about
operational plausibility and order-of-magnitude agreement, but not yet a
comprehensive statistical characterisation of model error across deployment
conditions.


\subsection{Model Derivation and Parametrisation}

Using the benchmark setup from Section~\ref{sec:measurement} and the registry generator described in Section~\ref{sec:methods}, we convert the Green Metrics Tool export into the JSON registry shown in Listing~\ref{lst:json-registry}. This demonstrates that the benchmark data is sufficient to instantiate all three model families introduced in Section~\ref{sec:methods}. In the resulting registry, \texttt{/login}, \texttt{/logout}, \texttt{/done}, and \texttt{/deleteAllToDos} are represented by constant models; \texttt{/createToDo} and \texttt{/getToDos} by piecewise curve models over runtime-visible payload proxies; and \texttt{/ai} by a token-linear model.

A key modelling choice concerns whether per-request energy is derived from idle-subtracted measurements or from total machine energy observed during the benchmark interval. Idle subtraction isolates the request-dependent signal, while total energy includes baseline consumption that may be relevant for provider-side accounting. In this work, we use total machine energy for model derivation and treat this as an explicit accounting assumption. Deployments using the method should therefore disclose whether reported values are marginal, average, or include baseline contributions.

The conversion step also yields deployment-specific carbon parameters shared across routes. For the measured system, grid intensity is \(\num{121.111}\,\mathrm{gCO_2e/kWh}\) and the embodied allocation rate is approximately \(\num{0.001}\,\mathrm{gCO_2e/s}\). These values are not universal constants but reflect the measured environment. Exposing them alongside request-level estimates ensures that the resulting disclosures remain interpretable.

\begin{table}[h]
\small
\centering
\caption{Endpoint benchmark conditions and derived per-request energy labels.}
\label{tab:energy}
\begin{tabular}{p{4.3cm}rr}
\toprule
\textbf{Endpoint / condition} & \textbf{Requests} & \textbf{\makecell[r]{Energy\\ {\footnotesize (mWh/req)}}} \\
\midrule
\texttt{/login}                              & 100 & 4.038 \\
\texttt{/logout}                             & 100 & 0.006 \\
\texttt{/done}                               & 100 & 0.085 \\
\texttt{/deleteAllToDos}                     & 100 & 2.835 \\
\midrule
\texttt{/getToDos} (100 items)               & 100 & 0.072 \\
\texttt{/getToDos} (1\,000 items)            & 100 & 0.112 \\
\texttt{/getToDos} (10\,000 items)           & 100 & 0.125 \\
\texttt{/getToDos} (100\,000 items)          & 100 & 0.553 \\
\texttt{/createToDo} (text, 100 chars)       & 100 & 0.064 \\
\texttt{/createToDo} (text, 1\,000 chars)    & 100 & 0.079 \\
\texttt{/createToDo} (text, 10\,000 chars)   & 100 & 0.035 \\
\texttt{/createToDo} (text, 100\,000 chars)  & 100 & 0.096 \\
\texttt{/createToDo} (file, 1\,KB)           & 100 & 0.058 \\
\texttt{/createToDo} (file, 10\,KB)          & 100 & 0.061 \\
\texttt{/createToDo} (file, 100\,KB)         & 100 & 0.080 \\
\texttt{/createToDo} (file, 1\,MB)           & 100 & 0.288 \\
\texttt{/createToDo} (file, 5\,MB)           & 100 & 0.989 \\
\midrule
\texttt{/ai} (very short prompt)             &  10 & 29.978 \\
\texttt{/ai} (short prompt)                  &  10 & 251.044 \\
\texttt{/ai} (medium prompt)                 &  10 & 299.771 \\
\texttt{/ai} (long prompt)                   &  10 & 229.718 \\
\texttt{/ai} (very long prompt)              &  10 & 274.852 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Endpoint behaviour}
We evaluate the fitted models across different endpoint classes to assess whether they preserve the expected workload-dependent behaviour. In particular, we distinguish between conventional CRUD-style endpoints and AI inference workloads, which exhibit fundamentally different scaling characteristics.

\subsubsection{CRUD endpoints}
The generated registry shows that the REST endpoints fall into distinct behavioural classes. Authentication-related requests are close to constant cost. The fitted \texttt{/logout} model is only \SI{0.006}{mWh} per request, while \texttt{/login} is substantially larger at \SI{4.038}{mWh}, reflecting additional work such as credential verification and cryptographic session establishment. The \texttt{/done} endpoint is also effectively constant, with a fitted value of \SI{0.085}{mWh} per request across all measured conditions.

In contrast, the payload-dependent endpoints exhibit clear scaling behaviour. For \texttt{/createToDo}, the fitted curve remains nearly flat for small text-only requests but increases sharply once file uploads become large. A request with a short text body requires about \SI{0.064}{mWh}, while the \SI{5}{MB} attachment condition reaches \SI{0.989}{mWh}, corresponding to more than a fifteen-fold increase. Intermediate file sizes (\SI{1}{KB}, \SI{10}{KB}, \SI{100}{KB}, \SI{1}{MB}) produce a consistent progression, indicating that the curve model captures a meaningful size effect rather than measurement noise.

For \texttt{/getToDos}, the fitted model scales with the size of the returned collection. The smallest measured response proxy is approximately \SI{49.4}{kB} and corresponds to \SI{0.072}{mWh} per request, while the largest condition reaches about \SI{20.3}{MB} and \SI{0.553}{mWh}. This shows that response-size observables available at the HTTP boundary are sufficient to approximate workload-dependent behaviour without access to internal database operations.

These results illustrate that the fitted models capture the dominant differences between endpoint classes while remaining compact. The goal is not to reproduce the full internal behaviour of each endpoint, but to provide interface-level abstractions that preserve the major decision-relevant distinctions between small and large requests.

\subsubsection{AI endpoint}
The AI endpoint highlights the importance of application-supplied variables. When considering only prompt-profile labels (\texttt{very-short}, \texttt{short}, \texttt{medium}, \texttt{long}, \texttt{very-long}), the observed runtimes are not strictly monotonic. In our measurements, the \texttt{medium} profile took longer than the \texttt{very-long} profile, which appears inconsistent with the expectation that longer prompts should require more work.

This behaviour is explained by the underlying token statistics. While prompt-token counts increase monotonically across profiles (from roughly 20 to about 1010 tokens), generated-token counts vary substantially and do not follow the same ordering. In this measurement, average generated-token counts were approximately 46.4 (\texttt{very-short}), 505.8 (\texttt{short}), 600.7 (\texttt{medium}), 446.5 (\texttt{long}), and 518.4 (\texttt{very-long}). As a result, prompt categories alone are not reliable predictors of runtime or energy.

Using token-level information, the fitted model for \texttt{/ai} is given by
\[
\widehat{E}_{ai}(\tau^{prompt}, \tau^{gen}) =
7.062 + 0.017 \cdot \tau^{prompt} + 0.483 \cdot \tau^{gen}.
\]
The coefficients make the relative contribution of the two factors explicit. The weight assigned to generated tokens is roughly 28 times larger than that of prompt tokens, indicating that the dominant marginal work occurs during token generation rather than prompt processing. This is consistent with the behaviour of decoder-style inference systems.

This result illustrates a key point of the paper: accurate per-request disclosure requires semantically meaningful application-level features. Generic HTTP-level observables such as payload size or prompt labels are insufficient to capture the dominant cost drivers of AI inference workloads.

\subsection{Click-Path Validation}
To assess how well the fitted endpoint models generalise beyond the isolated benchmark conditions used for calibration, we executed an additional no-header validation run that repeated a realistic user click-through path 100 times\footnote{https://metrics.green-coding.io/stats.html?id=fca8c396-7591-4f16-9fd5-c3f87c2a424f}. Each iteration performed the sequence \texttt{/login} \(\rightarrow\) \texttt{/deleteAllToDos} \(\rightarrow\) \texttt{/createToDo} \(\rightarrow\) \texttt{/getToDos} \(\rightarrow\) \texttt{/done} \(\rightarrow\) \texttt{/getToDos} \(\rightarrow\) \texttt{/logout}, resulting in 700 HTTP requests in total. The total run was measured ten times with Green Metrics Tool to obtain the machine-level reference energy and, independently, nginx access logs were replayed offline through the derived registry to obtain the model-based estimate.

The GMT reported a total machine energy of
\SI{597.18}{mWh} (\(\pm\)\,\SI{0.74}{\percent}, average \(\pm\) standard deviation over all runs)\footnote{\url{https://tinyurl.com/5t42rame}}, while the offline log replay produced an estimated total of \SI{611.101}{mWh}. This corresponds to an absolute deviation of
\SI{13.921}{mWh}, or \SI{2.33}{\percent} relative to the measured total. For an
endpoint model derived from separate controlled measurements and then evaluated
only from runtime-visible log features, this level of agreement is encouraging:
the estimate preserves the correct order of magnitude and remains close enough
to support relative comparisons and aggregate observability.

\subsection{Nginx Plugin Overhead}
We evaluate the runtime overhead of the disclosure mechanism using ten repeated-run comparisons with the nginx module disabled\footnote{\url{https://tinyurl.com/3he27b8w}} and enabled\footnote{\url{https://tinyurl.com/2rrn7bfb}}, while keeping the remaining setup unchanged. Across the complete benchmark run, the compare export reports mean machine energy increasing from $5325.07 \pm 0.48\% \mathrm{mWh}$ to $5349.55 \pm 0.50\% \mathrm{mWh}$, corresponding to an overall overhead of \SI{0.46}{\percent}. This aggregate value is derived from the full-run \texttt{[RUNTIME]} entry in the comparison export and is not repeated in the per-endpoint tables below.

At route level, the current data shows small effects for the conventional web endpoints (Tables~\ref{tab:compare-constant-endpoints}--\ref{tab:compare-file-loads}). In Table~\ref{tab:compare-constant-endpoints}, \texttt{/login} does not show a measurable change; the mean energy changes from $678.59 \pm 0.88\% \mathrm{mWh}$ to $676.31 \pm 0.95\% \mathrm{mWh}$, which is well within the observed run-to-run variation. The same table shows \texttt{/logout} increasing by \SI{10.9}{\percent}, but at only about \SI{15}{mWh} per block and with standard deviations around \SI{18}{\percent}, this is better interpreted as noise than as a stable overhead signal. In Table~\ref{tab:compare-text-loads}, the per-condition differences remain modest and mixed: \texttt{/getToDos} ranges from \SI{-3.8}{\percent} to \SI{+2.9}{\percent}, and \texttt{/done} from \SI{-4.2}{\percent} to \SI{+10.1}{\percent}. For \texttt{/createToDo}, Table~\ref{tab:compare-text-loads} shows text-condition changes between \SI{+4.0}{\percent} and \SI{+8.9}{\percent}, while Table~\ref{tab:compare-file-loads} shows file-upload changes between \SI{+0.1}{\percent} and \SI{+11.1}{\percent}. The largest relative increases occur on small absolute benchmark blocks, whereas the larger and more practically relevant upload cases (\SI{1}{MB} and \SI{5}{MB}) change by only \SI{0.2}{\percent} and \SI{0.1}{\percent}, respectively.

We do not include the AI endpoint in the overhead analysis below, because the repeated comparison runs did not yield a measurable overhead signal. Across the five prompt profiles, mean machine energy differences remained within \SI{-0.5}{\percent} to \SI{+0.7}{\percent}, and per-request runtime changes stayed between approximately \SI{0.0}{\percent} and \SI{0.3}{\percent}. Within this dataset, any effect of header emission remains below the natural variability of the inference workload.

Taken together, these results indicate that per-request model evaluation and disclosure can be integrated into the HTTP server with low overhead for both the conventional web routes studied here and, within the limits of the present comparison, the AI endpoint as well.

\begin{table}[h]
\centering
\scriptsize
\caption{Measured energy comparison for constant-cost authentication endpoints. Values show mean machine energy in mWh for each benchmark block.}
\label{tab:compare-constant-endpoints}
\begin{tabular}{llrr}
\toprule
Class & Variant & Login & Logout \\
\midrule
Constant & No Header & 678.59 $\pm$ 0.88\% & 14.42 $\pm$ 18.44\% \\
 & Header & 676.31 $\pm$ 0.95\% & 15.99 $\pm$ 17.26\% \\
 & $\Delta$ & -0.3\% & +10.9\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\scriptsize
\centering
\caption{Measured energy comparison for text-based workload steps. Values show mean machine energy in mWh for each benchmark block.}
\label{tab:compare-text-loads}
\begin{tabular}{llrrr}
\toprule
Load & Variant & CreateToDo & GetToDos & Done \\
\midrule
100 & No Header & 37.68 $\pm$ 5.74\% & 20.99 $\pm$ 5.49\% & 34.35 $\pm$ 14.57\% \\
 & Header & 39.33 $\pm$ 8.94\% & 20.21 $\pm$ 7.29\% & 37.82 $\pm$ 9.90\% \\
 & $\Delta$ & +4.4\% & -3.8\% & +10.1\% \\
\midrule
1000 & No Header & 39.63 $\pm$ 15.15\% & 21.31 $\pm$ 7.39\% & 34.95 $\pm$ 12.54\% \\
 & Header & 43.16 $\pm$ 10.45\% & 21.71 $\pm$ 6.67\% & 33.49 $\pm$ 13.69\% \\
 & $\Delta$ & +8.9\% & +1.9\% & -4.2\% \\
\midrule
10000 & No Header & 42.48 $\pm$ 10.15\% & 30.26 $\pm$ 4.15\% & 39.41 $\pm$ 9.45\% \\
 & Header & 44.20 $\pm$ 7.06\% & 31.14 $\pm$ 4.23\% & 38.84 $\pm$ 5.06\% \\
 & $\Delta$ & +4.0\% & +2.9\% & -1.5\% \\
\midrule
100000 & No Header & 46.05 $\pm$ 7.32\% & 131.92 $\pm$ 2.37\% & 45.58 $\pm$ 10.27\% \\
 & Header & 48.82 $\pm$ 7.45\% & 131.35 $\pm$ 1.90\% & 49.35 $\pm$ 8.72\% \\
 & $\Delta$ & +6.0\% & -0.4\% & +8.3\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\scriptsize
\centering
\caption{Measured energy comparison for file-upload workload steps. Values show mean machine energy in mWh for each benchmark block.}
\label{tab:compare-file-loads}
\begin{tabular}{llr}
\toprule
File size & Variant & CreateToDo \\
\midrule
1KB & No Header & 39.98 $\pm$ 10.67\% \\
 & Header & 41.42 $\pm$ 15.37\% \\
 & $\Delta$ & +3.6\% \\
\midrule
10KB & No Header & 36.27 $\pm$ 9.13\% \\
 & Header & 40.32 $\pm$ 10.89\% \\
 & $\Delta$ & +11.1\% \\
\midrule
100KB & No Header & 41.10 $\pm$ 10.31\% \\
 & Header & 44.73 $\pm$ 10.08\% \\
 & $\Delta$ & +8.8\% \\
\midrule
1MB & No Header & 68.66 $\pm$ 3.17\% \\
 & Header & 68.78 $\pm$ 3.64\% \\
 & $\Delta$ & +0.2\% \\
\midrule
5MB & No Header & 232.39 $\pm$ 0.80\% \\
 & Header & 232.62 $\pm$ 0.97\% \\
 & $\Delta$ & +0.1\% \\
\bottomrule
\end{tabular}
\end{table}


\section{Discussion}

The proposed approach is useful not only for external disclosure, but also for
internal provider-side analytics. Today, software producers typically have
access to request counts, latency, and cost, but lack visibility into the
energy or emissions associated with their own endpoints, as hosting providers
rarely expose fine-grained power data. By benchmarking services offline and
deriving per-endpoint models, providers can combine these models with routine
HTTP logs (e.g., request time, payload size, response size, and
application-level variables such as token counts) to estimate and aggregate
energy or \(\mathrm{CO_2e}\) across endpoints, customers, or time windows. This
provides a practical form of energy observability even when the underlying
infrastructure remains opaque.


At the same time, the resulting estimates depend on modelling assumptions and system boundaries. A single request may involve multiple components such as caches, databases, queues, or external services. In our implementation, the system boundary is defined by the benchmarked deployment, which can include co-measured dependencies but cannot account for opaque third-party services unless their impact is provided separately. Similarly, caching and intermediaries such as CDNs may shift where energy is expended, raising questions about how cache hits should be attributed (e.g., origin versus edge). Another important choice concerns whether reported values reflect marginal energy above idle or include a share of baseline system energy. Both interpretations are useful for different purposes, but must be made explicit for the resulting values to be interpretable.

More generally, the approach produces model-based estimates rather than direct measurements. The fitted models are calibrated on a reference system and may therefore differ from production deployments due to variations in hardware, virtualization, co-located load, or regional grid conditions. As a result, the disclosed values should be interpreted as context-dependent estimates rather than hardware-independent ground truth. In practice, this implies that models may require recalibration when deployment conditions change, and that their primary value lies in capturing relative differences between request classes rather than absolute precision.

The case for disclosure as a precondition for constraint rests on an
asymmetry: in current HTTP-based systems, clients and operators can
observe latency, cost, and throughput, but not environmental cost.
This asymmetry is not neutral.  It systematically excludes ecological
considerations from the concrete software decisions that collectively determine aggregate resource
consumption.  Making per-request impact visible does not, by itself,
reduce demand.  But it removes a structural barrier to doing so: one
cannot budget, cap, or differentially price what one cannot measure at
the relevant granularity.

At the same time, we recognise three risks.  First, per-request values
may be instrumentalised for greenwashing: a provider could disclose
small per-request numbers while obscuring that aggregate
demand continues to grow unchecked. Disclosure of marginal cost without disclosure of total volume provides
an incomplete and potentially misleading picture.  Second, the
mechanism could shift responsibility asymmetrically: if providers
disclose per-request impact but frame reduction as a client-side
concern (e.g., ``reduce your prompt length''), the burden of adaptation
falls on users while infrastructure growth remains unquestioned.
Third, the act of quantification itself can naturalise the assumption
that environmental impact is acceptable as long as it is accounted for.

Finally, the approach does not explicitly quantify model error or uncertainty. Measurement noise, modelling choices, and unobserved factors can all influence the resulting estimates. While this limits the interpretability of absolute values, it does not invalidate their usefulness. Similar to other interface-level metrics such as latency or throughput, the estimates remain valuable as long as they provide stable and decision-relevant signals. The key contribution is therefore not exact metrological accuracy, but a reproducible and inspectable method for making the environmental implications of software visible at the interface where decisions are made.

\section{Conclusion}
This paper argued that the main missing piece in HTTP sustainability disclosure is not only header-field semantics, but a practical method for generating credible per-request values. We showed that controlled machine-level measurements can be transformed into compact endpoint models, stored in a JSON registry, and evaluated online at the HTTP boundary by an nginx-based prototype. The evaluation demonstrates that this approach can represent heterogeneous endpoint classes, including both conventional CRUD routes and an AI inference endpoint, while remaining operationally simple enough for reverse-proxy deployment.

The current prototype still leaves room for systems-level optimisation. Although the measured overhead on the conventional web routes is small, it is not zero. An important next step is therefore to re-implement the runtime header-generation path as a native nginx extension in C. Doing so would remove the remaining JavaScript interpretation overhead, reduce per-request allocations further, and allow tighter integration with nginx's internal request-processing structures.

\section*{Availability}
All source code for the prototype, benchmarks, and supporting components is
freely available at \url{https://github.com/ribalba/Limits-2026/tree/main/code}.
All measurement runs referenced in this work are available at
\url{https://metrics.green-coding.io/runs.html}.


\bibliographystyle{ACM-Reference-Format}
\bibliography{bib}

\end{document}
